\begin{agradecimentos}
I would like to thank all my familly, but at first place my parents Joaquim Cândido and Édina Lúcia that are the main responsible for this achievement.  You have no idea the respect, admiration and love I have for you. You are my heros! Also my parents in law, Antônio and Arlete, precious gifts that life gave me that I keep inside my heart. 

My grandfathers Olavo Cândido and Otávio Perpétuo and my grandmothers Maria da Conceição (Lia) and Maria Salomé (Nina), \textit{in memoriam}.

My precious jewel of light, my wife Nikaelle Fraga, I'm really devoted to you! You suffer with me each step of this odyssey, each pain and each victory. You deserve more credits for this than me, for having supported me and encouraged me to follow my dreams. You give meaning to my favorite quote: \textit{Ubi thesaurus tuus est, ibi cor tuum est}. I love you!

My brothers André and Ivana, my best friends since ever and forever and people that always inspired me. All my uncles and cousins that I will not nominate to avoid any omission, also my sister-in-law Ingrid and my godchildren João Rafael and Antônio Nicolau. 

My advisor Prof. Dr. Frederico Gadelha Guimarães, the best advisor ever and really a gifted person. I'm really proud to be your student and it's a honor to have the opportunity to learn with you! Best wishes also for his wife Rúbia Pereira and his lovely daughter Laura.

My co-advisor Prof. Dr. Hossein Javedani Sadaei, a fantastic person which to a person who helped me and taught me a lot, also possibly the world's greatest specialist on Fuzzy Time Series. It's a privilege to count on you!

My lifelong best friends Prof. Dr. Gefter Thiago, Prof. Dr. Márcio Ramos (Bisa) and Prof. Rodrigo Carneiro Brandão. Even living so far and been so unplugged, my thoughts will always be with you! You are the best guys ever! 

All professors and staff of the PPGEE - UFMG and my teammates on MINDS - Machine Intelligence and Data Science Laboratory, very special and careful people! Marcos Alves, Carlos Severiano, Gustavo Vieira, Tamires Rezende, Rodrigo Pedrosa, Cristiano Leite, Ivan Reinaldo, Rúbia Reis, Leonardo Augusto,  Giulia Zanon, Maria Victória, Fernando Galindres, Roozbeh, Kossar, Omid Orang, Babak and Bruno Alberto. Thank you very much!  

All colleagues and students of IFNMG - Instituto Federal do Norte de Minas Gerais. I will not nominate to avoid any  omission, except for these special guys: Paulo Vitor (PV),  Felipe (Bolinha) and Patrícia Lucas, my colleagues at PPGEE.

Finally, my kittens Perrengo, Yôda, Salém and Vingador, my companies during the long nights writing this work. Meow!

\vspace{5cm}


\begin{center}
\textbf{\textit{Glória in Excélsis Deo!}}
\end{center}

\end{agradecimentos}